Examination of shapes! (homework exam)

### 2016-02-27 geometricAlgorithms/homeworkExam/homeworkExam.tex
running time and status structure added

After the border is found we pass it to the FindXYSilhouette function.
This funtion will take at most $O(N)$ time, because the function
istelf will always try to stay at much on the border as possible with help
of the least rotating edge function. This function traverses all connecting
edges. This sounds like a loop in a loop but it isn't really, because the
first loop doesn't consider all elements (only the outer ring) and the second
loop considers the connection in the outer rings. So some elements may be
considered dubble but this is just one element (the least rotating edge).
Its also worth noting that the ``inside'' edges may never be considered by
either of these functions because they try to stay on the outside.
\section{Red and blue lines}
\begin{algorithmic}[1]
	\State
	$P_{ends} \gets R \cup B$ (flat) Mapped to endpoints where each segment
	produces two end points, the endpoints contain a reference to eachother
	and they aslo contain which color they are and are also marked
	valid.
	\State
	$P_{sorted} \gets P_{ends}$ sorted on Y coordinate.
	\State
	$S_{tatus} \gets $ empty binary tree that sorts on X axis
	$S_{result} \gets $ empty list
	\ForAll{$p_{current} \gets Next(P_{sorted})$}
		\If{$p_{current}$ is a start point of a blue line}
			Insert $p_{current}$ into $S_{tatus}$
		\State
		\ElsIf{$p_{current}$ is a start point of a red line}
			\State
			$t_{neigbours} \gets $ Insert $p_{current}$ into
			$S_{tatus}$ which returns the left and right neigbours
			in a tupple
			\If{$t_{neigbours}$ are not both blue lines}
				\State
				$p_{current}$ gets marked invalid.
			\EndIf
		\State
		\ElsIf{$p_{current}$ is an end point of a blue line}
			\State
			$t_{neigbours} \gets $ Insert $p_{current}$ into
			$S_{tatus}$ which returns the left and right neigbours
			in a tupple
			\If{$t_{neigbours}$ a red line}
				\State
				mark that line invalid if $t_{neigbours}$ does
				not contain a blue line.
			\EndIf
			\State
			Remove $p_{current}$ and its twin (so the start point) from$S_{tatus}$
		\State
		\ElsIf{$p_{current}$ is an end point of a red line}
			\State
			$t_{neigbours} \gets $ Insert $p_{current}$ into
			$S_{tatus}$ which returns the left and right neigbours
			in a tupple
			\If{both $t_{neigbours}$ are blue and the twin of $p_{current}$ was still marked valid}
				Add $p_{current}$ to $S_{result}$
			\EndIf
			\State
			Remove $p_{current}$ and its twin (so the start point) from$S_{tatus}$
		\EndIf
	\EndFor
\end{algorithmic}

The status of the sweep line is a set of linesegments intersecting it at
its current position. ordered from left to right (x axis).

The events are the beginning of either a red or blue line and the ending of
a red or blue line.

@@

If the ending of a red line is encounred we check if its invalid,
if so we remove it from the structure and continue.
If its not invalid we insert the ending into the status structure.
If this ending is has a blue two blue lines as neigbors we add it
to the result set.
Finally we remove it and continue.
\section{Polygon cut}
\section{Diagonal polygon cut}
\end{document}

>>>>

After the border is found we pass it to the FindXYSilhouette function.
This funtion will take at most $O(N)$ time, because the function
istelf will always try to stay at much on the border as possible with help
of the least rotating edge function. This function traverses all connecting
edges. This sounds like a loop in a loop but it isn't really, because the
first loop doesn't consider all elements (only the outer ring) and the second
loop considers the connection in the outer rings. So some elements may be
considered dubble but this is just one element (the least rotating edge).
Its also worth noting that the ``inside'' edges may never be considered by
either of these functions because they try to stay on the outside.
\section{Red and blue lines}
\subsection{algoirthm}
To design the algorithm I wrote it down in psuedo code:
\begin{algorithmic}[1]
	\State
	$P_{ends} \gets R \cup B$ (flat) Mapped to endpoints where each segment
	produces two end points, the endpoints contain a reference to eachother
	and they aslo contain which color they are and are also marked
	valid.
	\State
	$P_{sorted} \gets P_{ends}$ sorted on Y coordinate.
	\State
	$S_{tatus} \gets $ empty binary tree that sorts on X axis, it also
	returns the two direct neigbours on insert.
	\State
	$S_{result} \gets $ empty list
	\For{$p_{current} \gets Next(P_{sorted})$}
		\If{$p_{current}$ is a start point of a blue line}
			Insert $p_{current}$ into $S_{tatus}$
		\State
		\ElsIf{$p_{current}$ is a start point of a red line}
			\State
			$t_{neigbours} \gets $ Insert $p_{current}$ into $S_{tatus}$

			\If{$t_{neigbours}$ are not both blue lines}
				\State
				$p_{current}$ gets marked invalid.
			\EndIf
		\State
		\ElsIf{$p_{current}$ is an end point of a blue line}
			\State
			$t_{neigbours} \gets $ Insert $p_{current}$ into
			$S_{tatus}$
			\If{$t_{neigbours}$ a red line}
				\State
				mark that line invalid if $t_{neigbours}$ does
				not contain a blue line.
			\EndIf
			\State
			Remove $p_{current}$ and its twin (so the start point) from$S_{tatus}$
		\State
		\ElsIf{$p_{current}$ is an end point of a red line}
			\State
			$t_{neigbours} \gets $ Insert $p_{current}$ into
			$S_{tatus}$
			\If{both $t_{neigbours}$ are blue and the twin of $p_{current}$ was still marked valid}
				Add $p_{current}$ to $S_{result}$
			\EndIf
			\State
			Remove $p_{current}$ and its twin (so the start point) from$S_{tatus}$
		\EndIf
	\EndFor
	\State
	\Return $S_{result}$
\end{algorithmic}

\subsection{Status}
The status of the sweep line is a set of linesegments intersecting it at
its current position. ordered from left to right (x axis).

The status is a binary tree with a special insert operation that also
returns the neigbours of the inserted element.

\subsubsection{Structure}
A node in the binary tree has a reference to a parent, and a left and right child.
A node has a value, which will be used to decide if you have to go left or right.
The initials node value is just the x coordinate.

\subsubsection{Insertion}
If another element gets
inserted we compare its x coordinate with that of the current node. If its
smaller we go left otherwise right. If x coordinate is equal we use the other
endpoint's X coordinate as a tie breaker, if that is equal to we look at
Y coordinate.
In case we are at the bottom of the tree we create a new connection node
and take the average of the old bottom element and the inserted element.
The parent node gets updated to let the right child point at the new
connection node and the new connection node has as children the old bottom
element and the newly inserted element.

\paragraph{Neigbours on insert}
But we defined a special binary tree that will return the neigbours upon
insertion. So at this point we already found one of the neigbours, to find
the other one we have to go up again until we can go into the neigbour
direction and this direction is not equal to where we cam from and then go
all the way in the other direction.
So if we try to find the right neigbour we go up and try to go to the right
child, if this is the child where we came from we go up further and try again.
Until we find the root, which means there is no right neigbor, or until
we find a child that is not the direction where we came from, which means
we have to go left all the way down in the tree.

\subsubsection{Removal}
Removal is done by finding the node we want to remove, then going to its
parent parent and saying one the child you came from is equal to its neigbor

\subsubsection{Balancing}
To ensure $O(log N)$ acces time an AVL implementation
can be used. But I think describing that is somewhat out of the scope
of this assignment.

\subsection{Events}
The events are the beginning of either a red or blue line and the ending of
a red or blue line.

@@

If the ending of a red line is encountered we insert the ending into the status structure.
If this ending has two blue lines as neigbors and its twin point wasn't marked invalid.
to the result set.
Finally we remove it and continue.

\subsection{Running time}
To do the plane sweep we first sort the points on Y access. This means
we have at least already a running time of:

\[ O((n+m) log (n+m))\]

Now the question is can we stay below this for the rest of the algorithm?
For the algorithm we consider all events which can be expressed as:

\[ 2(n+m) \]

Because for every red $n$ line there are two events, namely the beginpoint and
end point of that line.
So we loop trough this and use a binary tree to detirmin if there are neigbours
or not.
let $k$ be the size of $S_{tatus}$.
Doing the test for neigbours takes at most:

\[2 log(k)\]

Assuming its a balanced tree. Its times 2 because one neigbour is found for
free on insertion, but for the other neigbor in worst case we have to
consider the entire tree.

So the final running time for the algorithm is:

\[O(2(n+m) 2log(k))\]

Plus the initilization:

\[O((n+m)log(n+m)+2(n+m) 2log(k))\]

\section{Polygon cut}
\section{Diagonal polygon cut}
\end{document}
